---
title: "Multiple data frames"
author: "Daphné PERRY"
format: 
  html: default
  pdf : default 
---

```{r}
here::i_am("git-101.Rproj")
library(here)
```

# Data loading

```{r}
#| message: false
library(ggplot2)
library(dplyr)
library(tidyr)
library(vroom)
```

```{r}
planets<-vroom(here("data", "ALL_PLANETS.csv"))
stars<-vroom(here("data", "Liststars.csv"))
```
```{r}
sun_to_earth<-33300
```


# Planet oriented analyses

We look for the planet that has the largest ratio btw its mass and the mass of its star. We can't do that using the planets data frame only, we need to join it w/the stars data frame. 

## Inner join

```{r}
planet_star<-inner_join(planets, stars, by=join_by(star_code))
```
The order of the join is'nt important in an inner join. The simplest join_by I can have is when I have a vari name shared by the dataframes I'm working w/(which is the case here).Now I have 15 varis (8 from planets & 7 from stars bcs already have tthe star code). 

So the join is said to be inner in the sens that if the key is missing, then the corresponding planet isn't included in the new dataframe. Cad in an inner join we keep only matches btw the 2 dataframes (according the the 'join_by' vari(s)). Once we've done that, we can then do whatever we want, for ex, compute ratio: 

```{r}
planet_star|>
  mutate(mass_ratio = `Mass (Sun)`/`relative mass (earth)`*sun_to_earth)|>
  slice_max(mass_ratio)|>
  select(P_NAME, star, mass_ratio)
```
## Semi & Anti join

To analyse missing matches, how can we find them in the data ? For that we have 2 operations that are not really join (bcs don't prod a new dataframe) but linked to it bcs give what would be prod by a join and then jsp
```{r}
planet_wout_star<-anti_join(planets, stars, by=join_by(star_code))
```
This gives us the planets without stars (therefore they aren't included in tthe planet_star data frame that we've created w/the innerjoin). It finds the object in the left (hte first dataframe) that doesn't have a match in the second (the right dataframe). 

```{r}
star_wout_planet<-anti_join(stars, planets, by=join_by(star_code))
```
->27 stars don't have a planet in our dataframe. 

Something less usefull: semijoin = does the same thing in reverse : gives the planets that has a star & the stras that have a planet but w/out including the data (remove the data associated to star). 
Usefull: semi join on star of the planet, bcs stars can have multiple planets.Then we just get the planets of the star, w/out including the data about the planet. 
```{r}
stars_w_planets <- semi_join(stars, planets, by=join_by(star_code))
```

So the anti join is usefull in general (point potential prbl is our data). 

We have `r nrow(planet_wout_star)` planets w/out a corresponding star in the database and `r nrow(star_wout_planet)` stars w/out planets described in the database. 


## Left & right join
3 diff outer join : 

- the left one will keep all the data in the 1st table (on the left) even if they don't have a match on the 2nd table. 
- the right one will keep all the data in the 2nd table (on the left) even if they don't have a match on the 1st table.
- the full one will keep everything, even if they don't have a match

```{r}
all_planets<-left_join(planets, stars, by = join_by(star_code))
all_stars<-right_join(planets, stars, by = join_by(star_code))
all_all<-full_join(planets, stars, by = join_by(star_code))
```

Lots of stars that don't have a planet -> that's why we have more obs in all_stars than in all_planets. 
Choose right/left inner join depending on the stats we want to run, what we wanna do w/ the data. 

# Star oriented analysis
We're gonna interpre the absence of planet to the fact that there is non associated to the star: if a star doesn't appear in the planet data frame, this means that this star has no associated habitable planet. 

To compute the distribution of the average nb of habitable planet per star: 
```{r}
planet_star |>
  summarise(n=n(), .by=star_code) |>
  ggplot(aes(x=n))+
  geom_bar()
```
Large nb of stars w/ 1 habitable planet. This is wrong bcs we are missing here the stars w/out planet. 
```{r}
all_stars |>
  summarise(n=n(), .by=star_code) |>
  ggplot(aes(x=n))+
  geom_bar()
```

Still nobody w/ 0 planet, even though we did now on the all_star dataset, which includes the stars w/out planet. 

To fix this: 
```{r}
all_stars|>
  summarise(n=sum(!is.na(PLANET_IDX)), .by=star_code)|>
  ggplot(aes(x=n))+
  geom_bar()
```

Alt solution: 
```{r}
all_stars|>
  summarise(n=n_distinct(PLANET_IDX, na.rm=TRUE), .by=star_code)|>
  ggplot(aes(x=n))+
  geom_bar()
```




